### 2022-01-15 content/box2.profiles.Rmd

---
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<!--# 
This document is created from box2.profiles.Rmd, please made edits to code generating the table there.

Edit the table contents here:https://docs.google.com/spreadsheets/d/1R0_NGhAWJ3cxGKQVh06eHKxHGMCN6x5aNL3_qR3nAI0/edit?usp=sharing 
-->

```{r include=FALSE}
library(googlesheets4)
library(tidyverse)
library(knitr)
library(pander)
```

```{r include=FALSE}
df <- read_sheet("1R0_NGhAWJ3cxGKQVh06eHKxHGMCN6x5aNL3_qR3nAI0")
```

```{r results='asis'}
set.alignment("left")
df %>% 
  replace_na(replace = list(Role = "", Description = "")) %>%
  pander(split.table = 10000, caption = "Collaborator profiles and some example use cases for GitHub features.  This is a non-exhaustive list intended to showcase some creative uses of GitHub for collaborative academic projects, such as collaborative manuscript writing.")
```