### 2022-04-19 R/comparison_table.Rmd

---
knit: (function(input_file, encoding) {
  out_dir <- '../content';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, '11.comparison_table.md'))})
output: 
  md_document:
    variant: gfm #forces table to be markdown, not html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
library(tidyverse) #data wrangling
library(here) #for reproducible relative file paths
library(flextable) #another possible table formatting package.  Can output to html, word, powerpoint, pdf.  Supports cell merging
options(knitr.table.format = 'pipe')
```

<!--DO NOT EDIT. Generated from R/comparison_table.Rmd, so edit that and render it to make changes.  This should get tacked on to the end of 09.figures-tables.md when compiled by Manubot -->

```{r wrangle, include=FALSE}
comp_tbl <- read_csv(here("data", "comparison_table_raw.csv"))
```

```{r compare}
kable(comp_tbl, caption = "Table: a comparison of technologies... {#tbl:compare}")
```

<!-- An example citation of this table (Table @tbl:compare). -->