### 2022-06-01 R/scatterblob.Rmd
fixing x and y axis labels

---
title: "Figure: the scatterblob"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(googlesheets4)
library(here)
library(readr)
```

Goal: This is a summary figure, e.g., "here are all the ways and here are the entry points with low technical difficulty, and some  for beginner s and the more advanced ways to use it".
There are ways to use github that allows different degrees of collaboration that do not require too many technical skills.

Use even notation across text and figure:
homogenizing titles of ways to use github

## The data

Data was generated using Google docs, then manually saved as csv file "scatterblob_data_raw.csv" in the `data` folder.

Data for this figure was generated on a collaborative [google spread sheet](https://docs.google.com/spreadsheets/d/1Ls0DofQWSZQsqtP4OLFj326sGfWsklrI2tA1GNUTR6g/edit#gid=0).

[Data, version 2](https://docs.google.com/spreadsheets/d/1zCRgctjfjE2xZN9MkqfFh222K6bvVvA3gT0UdkL_CS8/edit#gid=0).

Cognitive load [was measured in a different file](https://docs.google.com/spreadsheets/d/1OLBR3M0uV3PMSazWUYHgPeX28wllAgS4PIlJXaNu-2o/edit#gid=0).

[Cognitive load, version 2](https://docs.google.com/spreadsheets/d/1YipCCdg5Z8w-IQO67rFfMBSTpLjO-_7HiW44gqzTXPM/edit#gid=0).

Reading the data csv file:

```{r}
scatterblob_data <- read_csv(here("data", "scatterblob_data_raw.csv"))
#  View(scatterblob_data)
```

Creating the `my_data` object from :

```{r}
names(scatterblob_data)
scatterblob_data$x1 <- scatterblob_data$`Min degree of collaborativeness X1`
scatterblob_data$x2 <- scatterblob_data$`Max degree of collaborativeness X2`
scatterblob_data$y_real <- scatterblob_data$`Cognitive load - real - non-overlapping`
scatterblob_data$y_seq <- scatterblob_data$`Cognitive load - sequential`
my_data <- scatterblob_data[1:12,c("Ways to use GitHub","y_seq", "y_real", "x1", "x2")]
 my_data
```

## Plotting the data
```{r}
outpath <- here("content", "images")
```

### First try

Using original cognitive values.

```{r}
my_colors <- viridis::turbo(n = nrow(my_data), alpha = 0.9)

some_colors <- viridis::turbo(n = 30, alpha = 0.9)

plot(1:30, 1:30, col = some_colors, pch = 8)
```

```{r}
my_colors <- some_colors[c(3:4,
                           8,10,12,
                           20,21,22,23,24,
                           28,30)]
###############################################################
# start the pdf/png connection and set graphical parameters
###############################################################
plot_format <- "pdf"
# plot_format <- "svg"
# plot_format <- "png"

if (plot_format == "pdf") {
  pdf(file = here(outpath, "scatterblob_0.pdf"),
      width = 8.25,
      height = 6,
      bg = "white")  
}
if (plot_format == "svg") {
  svg(filename = here(outpath, "scatterblob_0.svg"),
      width = 8.25,
      height = 6,
      bg = "white")  
}
if (plot_format == "png") {
  png(file = here(outpath, "scatterblob_0.png"),
      width = 8.25,
      height = 6,
      units = "in",
      res = 100,
      bg = "white")  
}
par(xpd = NA,
    mai = c(1.02, 1, 0.82, 4),
    bty="n")
###############################################################
# create the plot background
###############################################################
plot(x = c(0.5, 5.5),
     y = c(1.75, 9.75),
     xlab = "",
     ylab = "",
     col = "white",
     # "hide" the tick labels so we can put some words instead of numbers:
     col.axis = "white",
     tck = 0.02)
title(ylab="Technical difficulty", line=4, cex.lab=1)
title(xlab="Degree of collaborativeness", line= 3.5, cex.lab=1)
###############################################################
# add tick labels
###############################################################
text(x = 1:5,
     y = c(1, rep(1.05, 4)),
     cex = 0.7,
     labels = c("Personal",
                "Lab \nmembers",
                "Other \nlabs",
                "The \ncommunity",
                "The \npublic"))
text(x = c(1, 3, 5),
     y = 0.5,
     labels = c("Low", "Medium", "High"))
# Technical difficulty levels from https://www.londonschool.com/level-scale/
text(x = 0.2,
     y = c(2, 4, 6, 8, 10),
     cex = 0.7,
     adj = 1,
     labels = c("Elementary",
                "Low \nintermediate",
                "Intermediate",
                "Pre \nAdvanced",
                "Very \nAdvanced"))
###############################################################
# plot scatter blobs
###############################################################
for (i in seq(nrow(my_data))) {
  X0 <- my_data$x1[i]
  X1 <- my_data$x2[i]
  Y <- my_data$y_real[i]
  if (X0 == X1) {
    X0 <- X0 - 0.01
    X1 <- X1 + 0.01
   }
  segments(x0 = X0,
           x1 = X1,
           y0 = Y,
           col = my_colors[i],
           lwd = 13)
}
###############################################################
# add github uses text
###############################################################
text(x = 5.5,
    y = my_data$y_real,
    labels = my_data$`Ways to use GitHub`,
    adj = 0,
    col = my_colors)
dev.off()
```
![Figure 1. Degree of collaboration VS technical difficulty](scatterblob_0.pdf)

### Second try

Distributing cognitive load evenly:

```{r}
some_colors <- viridis::turbo(n = 100, alpha = 0.9)

plot(1:100, 1:100, col = some_colors, pch = 8)
```

```{r}
my_colors <- some_colors[c(15,20,
                           26,28,30,
                           66,68,70,72,74,
                           85,95)]
###############################################################
# start the pdf/png/svg connection and set graphical parameters
###############################################################
#plot_format <- "pdf"
# plot_format <- "png"
 plot_format <- "svg"

if (plot_format == "pdf") {
  pdf(file = here(outpath, "scatterblob_1.pdf"),
      width = 8.25,
      height = 6,
      bg = "white")  
}
if (plot_format == "svg") {
  svg(filename = here(outpath, "scatterblob_1.svg"),
      width = 8.25,
      height = 6,
      bg = "white")  
}
if (plot_format == "png") {
  png(file = here(outpath, "scatterblob_1.png"),
      width = 8.25,
      height = 6,
      units = "in",
      res = 100,
      bg = "white")  
}
par(xpd = NA,
    mai = c(1.02, 1, 0.82, 4),
    bty="n")
###############################################################
# create the plot background
###############################################################
plot(x = c(0.5, 5.5),
     y = c(0.4, 12.5),
     xlab = "",
     ylab = "",
     col = "white",
     # "hide" the tick labels so we can put some words instead of numbers:
     col.axis = "white",
     tck = 0.02)
title(ylab="Technical difficulty", line=4, cex.lab=1)
title(xlab="Degree of collaborativeness", line= 3.5, cex.lab=1)
###############################################################
# add tick labels
###############################################################
text(x = 1:5,
     y = c(-0.65, rep(-0.7, 4)), 
     cex = 0.7,
     labels = c("Personal", 
                "Lab \nmembers", 
                "Other \nlabs", 
                "The \ncommunity", 
                "The \npublic"))
text(x = c(1, 3, 5),
     y = -1.7, 
     labels = c("Low", "Medium", "High"))
# Technical difficulty levels from https://www.londonschool.com/level-scale/
text(x = 0.2,
     y = c(0, 1.5, 3, 4.5, 6, 7.5, 9, 10.5, 12 ), 
     cex = 0.7,
     adj = 1,
     labels = c("Beginner",
                "Elementary", 
                "Low\nintermediate",
                "Pre\nintermediate",
                "Intermediate", 
                "Upper\nintermediate",
                "Pre\nAdvanced",
                "Advanced",
                "Very\nAdvanced"))
###############################################################
# plot scatter blobs on sequential cognitive load
###############################################################
for (i in seq(nrow(my_data))) {
  X0 <- my_data$x1[i]
  X1 <- my_data$x2[i]
  Y <- my_data$y_seq[i]
  if (X0 == X1) {
    X0 <- X0 - 0.01
    X1 <- X1 + 0.01
   }
  segments(x0 = X0,
           x1 = X1,
           y0 = Y,
           col = my_colors[i],
           lwd = 23)
}
###############################################################
# add github uses text
###############################################################
text(x = 5.5,
    y = my_data$y_seq,
    labels = my_data$`Ways to use GitHub`,
    adj = 0,
    col = my_colors)
dev.off()
```

![Figure 2. Degree of collaboration VS Even technical difficulty](scatterblob_1.pdf)