### 2023-03-09 R/comparison_table.Rmd
Fixing output_file for 10.comparison_table.md

---
knit: (function(input_file, encoding) {
  out_dir <- '../content';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, '10.comparison_table.md'))})
output: 
  md_document:
    variant: gfm #forces table to be markdown, not html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
library(tidyverse) #data wrangling
library(here) #for reproducible relative file paths
library(flextable) #another possible table formatting package.  Can output to html, word, powerpoint, pdf.  Supports cell merging
options(knitr.table.format = 'pipe')
```

<!--DO NOT EDIT. Generated from R/comparison_table.Rmd, so edit that and render it to make changes.  This should get tacked on to the end of 09.figures-tables.md when compiled by Manubot -->

```{r wrangle, include=FALSE}
comp_tbl <- read_csv(here("data", "comparison_table_raw.csv"))
```

```{r compare}
kable(comp_tbl, caption = "A comparison of technologies commonly used for collaborating on research in Ecology and Evolutionary Biology. In the first column, we group platforms for collaboration into broad guilds. The second column lists the platform for collaboration. The remaining columns indicate whether the platform for collaboration includes certain features {#tbl:compare}")
```

<!-- An example citation of this table (Table @tbl:compare). -->